---
title: "Session 08: Functions, Pipes and Iteration"
author: "326.212 전산통계 및 실습"
institute: "서울대학교 통계학과"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```


# Prerequisite

```{r}
library(tidyverse)
library(magrittr)
```


# Functions

## Why do we use functions?

Functions allow you to automate common tasks in a more powerful and general way than copy-and-pasting.

Advantages over using copy-and-paste:

* You can give a function an evocative name that makes your code easier to understand.

* As requirements change, you only need to update code in one place, instead of many.

* You eliminate the chance of making incidental mistakes when you copy and paste (i.e. updating a variable name in one place, but not in another).

## Components of a function

```{r}

# name and argument
is_odd <- function(x) {
  "
    Describe your function in here!
  "
  
  # checking for exception
  if (!is.numeric(x)) {
    stop("The input should be a numeric value")
  }
  
  if (length(x) != 1) {
    stop("The input should be a numeric value")
  }
  
  # body
  if (x %% 2 == 0) {
    return(FALSE)
  } else {
    return(TRUE)
  }
  
}

```

#### Names

* Clear and short name!

* Naming style: `snake_case` vs `camelCase`

* R allows 'dot' in names like `dont.do.this`

* Avoid overriding exisiting functions and variables

#### Arguments

Detail arguments should go on the end, and usually should have default values.

```{r}

# Compute confidence interval around mean using normal approximation
mean_ci <- function(x, conf = 0.95) {
  se <- sd(x) / sqrt(length(x))
  alpha <- 1 - conf
  mean(x) + se * qnorm(c(alpha / 2, 1 - alpha / 2))
}

x <- runif(100)
mean_ci(x)

```

**Do not** put default valued arguments before non-default valued arguments.

```{r}

# good
no_use_sum <- function(a, b, c = 0) {
  return(a+b+c)
}

# bad
no_use_sum <- function(a, b = 0, c) {
  return(a+b+c)
}

```

Do exception hadling. (Using `try()` and `tryCatch()` would be good choice. We also learned about `safely()`, `possibly()` and `quietly()`.)

```{r, eval = FALSE}

is_odd(1)

is_odd("1")

is_odd(c(1,2))

```

#### Body

The body of function is the main part that what you want to do with function will be programmed. Two main techniques - Conditions and Iterations - will be introduced later.

#### Return

The value returned by the function is usually the last statement it evaluates when you don't specify return of fucntion.

## Writing pipeable functions

Side-effects functions should “invisibly” return the first argument, so that while they’re not printed they can still be used in a pipeline:

```{r}

show_missings <- function(df) {
  n <- sum(is.na(df))
  cat("Missing values: ", n, "\n", sep = "")
  
  invisible(df)
}

```

```{r}

show_missings(mtcars)

```

```{r}

mtcars %>% 
  show_missings() %>% 
  mutate(mpg = ifelse(mpg < 20, NA, mpg)) %>% 
  show_missings() 

```

```{r}

show_missings2 <- function(df) {
  n <- sum(is.na(df))
  cat("Missing values: ", n, "\n", sep = "")
  
  df
}

mtcars %>% 
  show_missings2() %>% 
  mutate(mpg = ifelse(mpg < 20, NA, mpg)) %>% 
  show_missings2() 

```

## Environment

```{r}

f <- function(x) {
  x + y
}

y <- 100

f(10)

y <- 1000

f(10)

```

```{r}

x <- 10
y <- 20

f <- function() {
  x <- 100
  y <- 200
  
  x + y
}

f()

x

y

```

```{r, eval = FALSE}

rm(y)

x <- 10

f <- function() {
  x <- 100
  y <- 200
  
  x + y
}

f()

y

```

```{r}

x <- 10

f <- function() {
  # avoid using this allocation(super assignment)
  x <<- 100
  
  x
}

f()

x

```

## Conditions

`&`, `|`, `&&`, `||`

```{r}

c(TRUE,FALSE) & TRUE

```

```{r}

c(TRUE,FALSE) && TRUE

```

```{r}

c(TRUE,FALSE) | TRUE

```

```{r}

c(TRUE,FALSE) || TRUE

```

`all`, `any`

```{r}

all(c(TRUE,FALSE) & TRUE)

```

```{r}

any(c(TRUE,FALSE) & TRUE)

```

`==`, `identical`

```{r}

c(1, 3) == c(3, 3)

```

```{r}

identical(c(1, 3), c(3, 3))

```

`if_else()`

```{r}

multiple_op_ftn <- function(x, y, op) {
  if (op == 'plus') {
    x + y
  } else if (op == 'minus') {
    x - y
  } else if (op == 'times') {
    x * y
  } else if (op == 'divide') {
    x / y
  } else {
    stop("Unknown op!")
  }
}

multiple_op_ftn(c(2,4), 3, 'minus')

multiple_op_ftn(c(2,4), 3, 'times')

```

`switch`

```{r}

multiple_op_ftn <- function(x, y, op) {
  switch(op,
    plus = x + y,
    minus = x - y,
    times = x * y,
    divide = x / y,
    stop("Unknown op!")
  )
}

multiple_op_ftn(c(2,4), 3, 'minus')

multiple_op_ftn(c(2,4), 3, 'times')

```

`ifelse()`

```{r}

x <- 10

ifelse(is_odd(x), "I'm odd!", "I'm even...")

ifelse(is_odd(x + 1), "I'm odd!", "I'm even...")

```

```{r}

mtcars %>% mutate(
  cyl = ifelse(
    cyl == 4, "I'm 4", ifelse(
      cyl == 6, "I'm 6", "I'm 8"
    )
  )
)

```

`dplyr::case_when`

```{r}

mtcars %>% mutate(
  cyl = case_when(
    cyl == 4 ~ "I'm 4",
    cyl == 6 ~ "I'm 6",
    cyl == 8 ~ "I'm 8"
  )
)

```

# Pipes

`T pipe`

```{r}

1:10 %>% mean()

1:10 %T>% mean()

```

```{r}

rnorm(100) %>%
  matrix(ncol = 2) %T>%
  plot() %>%
  str()

```

`%$%`

```{r}

mtcars %$%
  cor(disp, mpg)

```

```{r}

cor(mtcars$disp, mtcars$mpg)

```

`%<>%`

```{r}

set.seed(1120)

x <- rnorm(10)

(x <- x %>% matrix(ncol = 2))

```

```{r}

set.seed(1120)

x <- rnorm(10)

(x %<>% matrix(ncol = 2))

```

# Iterations

## Imperative Styles: Loops

### Basic control flows

Taylor series of $e^x$ is:

$$e^x = \sum_{n=0}^\infty\dfrac{x^n}{n!}.$$

```{r}

my_exp <- function(x) {
  
  result = 0
  
  for (i in 0:100) {
    result = result + x^i / factorial(i)
  }
  
  result
}

my_exp(10)

exp(10)

```

```{r}

my_exp <- function(x) {
  
  result = 0
  
  i = 0
  
  while (i <= 100) {
    result = result + x^i / factorial(i)
    i = i + 1
  }
  
  result
}

my_exp(10)

exp(10)

```

`break`

```{r}

my_exp <- function(x) {
  
  result = 0
  
  i = 0
  
  while (1) {
    result = result + x^i / factorial(i)
    i = i + 1
    if (i > 100) {
      cat("Break occurs at ", i, "\n")
      break
    }
  }
  
  result
}

my_exp(10)

exp(10)

```

`next`

```{r}

x <- rnorm(10000)

y = 0

# sum all positive values
for (i in seq_along(x)) {
  if (x[i] <= 0) {
    next
  } else {
    y = y + x[i]
  }
}

y

sum(x[x > 0])

```

### Performance tip

** Do not ** generate your result as follows.

```{r}

Ns <- c(1000, 2000, 5000, 10000, 20000)
result <- vector("list", length(Ns)); names(result) <- Ns
for (idx in seq_along(Ns)) {
  my_seq <- 1:Ns[[idx]]
  res <- system.time({
    foo <- double()
    for (i in seq_along(my_seq)) {
      foo <- c(foo, rnorm(sample(10, 1), my_seq[[i]]))
    }
  })
  result[[idx]] <- res[3]
}
(result <- unlist(result))              # worse than O(n^2)

```

```{r}

result <- vector("list", length(Ns)); names(result) <- Ns
for (idx in seq_along(Ns)) {
  my_seq <- 1:Ns[[idx]]
  res <- system.time({
    bar <- vector("list", Ns[[idx]])
    for (i in seq_along(my_seq)) {
      bar[[i]] <- rnorm(sample(10, 1), my_seq[[i]])
    }
    bar <- unlist(bar)
  })
  result[[idx]] <- res[3]
}
(result <- unlist(result))              # O(n)

```

If you don't know length of output, generate `list`!

```{r}

means <- c(0, 1, 2)

out <- vector("list", length(means))
for (i in seq_along(means)) {
  n <- sample(100, 1)
  out[[i]] <- rnorm(n, means[[i]])
}

str(out)

```

```{r}

str(unlist(out))

```

## Functional Styles: map, walk, reduce and accumulate

```{r}

(foo <- tibble(a = c(1, 2), b = c(3, 4), c = c(5, 6)))

```

```{r}

sapply(foo, mean)                       # `*apply` in base R

```

```{r}

map_dbl(foo, mean)                      # `map_*` in purrr package

```

```{r}

map_dbl(foo, function(x) x[[2]]^x[[1]]) # using anonymous function

```

For more multiple argument, you can use `map2()`, `pmap()`, etc...

```{r}

x <- c(1, 7, 3, 4, 8, 10, 5, 2, 6, 9)
reduce(x, `+`); reduce(x, `+`, .dir = "backward")

```

```{r}

accumulate(x, `+`); accumulate(x, `+`, .dir = "backward")

```

## Vectorization

R is not fast language, even loops in R is painfully slow. For higher performance, use vectorized functions. If you know C++ language, `Rcpp` would be a candidate for boost-up your program.

```{r}

x <- rnorm(1000000) %>% matrix(1000)

y1 = vector('numeric', 1000)

system.time({
  for (i in 1:1000) {
    for (j in 1:1000) {
      y1[i] <- y1[i] + x[j,i]
    }
  }
})

```

```{r}

system.time({ y2 <- apply(x, 2, sum) })

```

```{r}

system.time({ y3 <- colSums(x) })

```

```{r}

all(near(y1, y2))

all(near(y2, y3))

```


-----

**Acknowledgment** All contents in this note is based on the book “R for Data Science”, written by Grolemund & Wickham.