---
title: "Session 06: Import and Tidy Data / Relational Data / Strings"
author: "326.212 전산통계 및 실습"
institute: "서울대학교 통계학과"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
rm(list = ls())
options(repos = c(CRAN = "http://cran.rstudio.com"))

library("ggplot2")
library("dplyr")
library("tidyr")
library("readr")
library("stringr")
```


# Import and Tidy Data

## Tibbles

```{r}
df_tibble <- tibble(
  x = runif(5),
  y = rnorm(5)
)
df_tibble
```

```{r}
df_tibble$x; df_tibble[["x"]]; df_tibble[[1]]; df_tibble %>% .$x
```

```{r}
df_dataframe <- as.data.frame(df_tibble)

df_dataframe[,2] # == df_tibble[[2]]
df_dataframe[,"y"] # == df_tibble[["x"]]
```

- You need to get used to both `data.frame` and `tibbles`

## `readr` package

- Load data

```{r}
money <- read_csv("lab_06_data_money.csv") # do not change this line
str(money)
```

```{r}
money
```


```{r}
money.without.parser <- read.csv("lab_06_data_money.csv")
str(money.without.parser)
```

```{r}
money.without.parser
```


- How to change each column as we expect? For example, try to make the following data frame;

```{r, echo = FALSE}
apply(money.without.parser, 2, parse_number)
```

## factor vs character

```{r}
fruit_ch <- fruit[1:5]
fruit_ch
class(fruit_ch)
```

```{r}
fruit_fa <- factor(fruit[1:5])
fruit_fa
class(fruit_fa)
```

```{r}
fc<-c("Kang", "Doo", "Young")
fc_f=as.factor(fc)
fc_f
```

```{r}
as.numeric(fruit_fa)
c(fruit_fa, "dragonfruit")
fruit_fa[1] <- "dragonfruit"
fruit_fa
```

```{r}
as.numeric(fruit_ch)
c(fruit_ch, "dragonfruit")
fruit_ch[1] <- "dragonfruit"
fruit_ch
```


### Suppress `factor` when reading data

- If you don't like `factor`, how to suppress it?

```{r, eval = FALSE}
money <- read.csv("lab_06_data_money.csv") # I don't like factor!!
```

```{r}
str(money) 
```

- Find an option in `read.csv` that suppresses the conversion into `factor`, and get the result like this;

```{r, echo = FALSE}
money <- read.csv("lab_06_data_money.csv", stringsAsFactors = F) # do not change this line
```

```{r}
str(money)

```


## Gathering

```{r}
table4a
```

```{r}
table4a %>%
  gather(`1999`, `2000`, key = "year", value = "cases")
```


## wide format vs long format
```{r}
table2 # long
```

```{r}
table2 %>% spread(key = year, value = count) # wide
```


## long format is preferred in ggplot
```{r}
ggplot(table2, aes(x = country, y = count, color = factor(year))) + geom_bar(stat = "identity") + facet_wrap(~ type, scales = "free")

```

## 12.6 Case study
```{r}
data(who)
head(who)
who %>%
  gather(key = "key", value = "value", new_sp_m014:newrel_f65, na.rm = TRUE) %>% 
  mutate(key = stringr::str_replace(string = key, pattern = "newrel", replacement = "new_rel")) %>%
  separate(col = key, into = c("new", "var", "sexage"), sep = "_") %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(col = sexage, into = c("sex", "age"), sep = 1)
```



# Relational Data

```{r}
library(nycflights13)
```

```{r}
flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier)
flights2
```


```{r}
flights2 %>% left_join(weather)
```

```{r}
planes
```


```{r}
flights2 %>%  left_join(planes, by = "tailnum")
```


```{r}
flights2 %>%  left_join(airports, c("dest" = "faa"))
```

```{r}
top_dest <- flights %>%
  count(dest, sort = TRUE) %>%
  head(10)
top_dest
```

```{r}
top_flights <- flights %>% semi_join(top_dest) 
top_flights %>% .$dest %>% unique
```

```{r}
flights %>% 
  anti_join(planes, by = "tailnum") %>%
  count(tailnum, sort = TRUE)
```

```{r}
flights %>% 
  anti_join(planes, by = "tailnum") %>% .$tailnum %>% unique
```

```{r}
setdiff(flights$tailnum, planes$tailnum)
```


# Strings


## Print something in loop
```{r, eval = FALSE}
for(i in 1:10){
  obj_print <- i
  cat(obj_print)
}
```


```{r, eval = FALSE}
for(i in 1:10){
  obj_print <- i
  obj_print
}
```

- Nothing happens... What should we do?

```{r, eval = FALSE}
for(i in 1:10){
  obj_print <- i
  print(obj_print)
}
```

## `sprintf`

- Do identically what C function `sprintf` does

- See an example below;

```{r, eval = FALSE}
gender <- "M" # or "F"
your_name <- "Seongoh Park"

sprintf("Current time is %s, %s. %s", 
        Sys.time(),
        if(gender=="M") "Mr" else "Ms",
        your_name)

# %03d
# %3.2f

# apply this to multiple cases
info <- data.frame(your_name = c("A", "B", "C"),
           gender = c("M", "M", "F"))
info
sprintf("Current time is %s, %s. %s", 
        Sys.time(),
        ifelse(info$gender =="M", "Mr", "Ms"),
        info$your_name)
```

- Other types are also available

```{r, eval = FALSE}
for(i in 1:5){
  cat(sprintf("integer i= %d\n", i))
  cat(sprintf("double i= %f\n", i))
  cat(sprintf("double i= %.3f\n\n", i))
}
```




## Combining strings

```{r}
str_c("x", "y", sep = ", ")
```

```{r}
str_c("prefix-", c("a", "b", "c"), "-suffix")
```

```{r}
str_c(c("x", "y", "z"), collapse = ", ")
```

## Matching patterns with regular expressions

```{r}
x <- c("abc", "a.c", "bef")
x
str_view(x, "a.")
```

```{r}
str_view(x, "a\\.c")  
```

- why "\\."?

```{r}
writeLines("a\\.c")
```

```{r}
bananas.is.good <- c("banana", "Banana", "BANANA")
str_view(bananas.is.good, "banana")
str_view(bananas.is.good, regex("banana",))
```

```{r}
str_view(bananas.is.good, regex("banana", ignore_case = TRUE))
```

```{r}
x <- "Line 1\nLine 2\nLine 3"
str_extract_all(x, "^Line")[[1]]
```

```{r}
str_extract_all(x, regex("^Line", multiline = TRUE))[[1]]
```
