---
title: "HW4"
author: "박찬영"
date: "2024-11-30"
output:
  rmarkdown::html_document:
    theme: lumen
---

<style>
  .answer{
    background-color: #f8f9fa;
    color: #343a40;
    font-family: 'Arial', sans-serif;
    font-size: 16px;
    line-height: 1.5;
    padding: 10px 15px;
    border-left: 4px solid #007bff;
    border-radius: 5px;
    margin: 15px auto;
    max-width: 800px;
    word-break: break-word;
  }
  </style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE,
  warning = FALSE,
  fig.width = 6,
  fig.height = 4,
  out.width = "70%",
  fig.align = "center"
)
```

```{r library, message = FALSE}
library(tidyverse)
```

### Textbook(1st) 19.2.1 problem 4

write your own functions to compute the variance and skewness of a numeric vector. <br>
Variance is defined as
$$
\text{Var}(x) = \frac{1}{n-1} \sum_{i=1}^n (x_i - \bar{x})^2
$$
where $\bar{x} = {\sum_{i=1}^n x_i}/n$ is the sample mean. <br>
Skewness is defined as
$$
\text{Skew}(x) = \frac{\frac{1}{n-2} \left( \sum_{i=1}^n (x_i - \bar{x})^3 \right)}{\text{Var}(x)^{3/2}}.
$$

```{r 1}
my_var <- function(x) {
  if (!is.numeric(x)) {
    stop("input numeric vector")
  }
  y <- na.omit(x)
  bary <- mean(y)
  vary <- sum((y - bary)^2) / (length(y) - 1)
  return(vary)
}

x <- rnorm(100)
var(x)
my_var(x)

my_skew <- function(x) {
  if (!is.numeric(x)) {
    stop("input numeric vector")
  }
  y <- na.omit(x)
  bary <- mean(y)
  u <- sum((y - bary)^3) / (length(y) - 2)
  vary <- sum((y - bary)^2) / (length(y) - 1)
  skew <- u / vary^(3 / 2)
  return(skew)
}
e1071::skewness(x)
my_skew(x)
```

### Textbook(1st) 19.2.1 problem 5

Write `both_na()`, a function that takes two vectors of the same length and returns the number of positions that have an `NA` in both vectors.

```{r 2}
both_na <- function(x, y) {
  if (length(x) != length(y)) {
    stop("길이가 같은 벡터를 입력하세요.")
  }
  na_x <- is.na(x)
  na_y <- is.na(y)
  logic <- na_x * na_y
  pos <- c()
  for (i in seq_len(length(x))){
    if (logic[i] == 1) {
      pos <- c(pos, i)
    }
  }
  return(pos)
}
both_na(c(1, NA, 2, 3, NA, 8, NA), c(NA, NA, 3, 4, NA, 8, 10))
```

### Textbook(1st) 19.4.4 problem 4

How could you use `cut()` to simplify this set of nested if-else statements?

```{r test, eval = FALSE}
if (temp <= 0) {
  "freezing"
} else if (temp <= 10) {
  "cold"
} else if (temp <= 20) {
  "cool"
} else if (temp <= 30) {
  "warm"
} else {
  "hot"
}
```
How would you change the call to `cut()` if I’d used `<` instead of `<=`? What is the other chief advantage of `cut()` for this problem? *(Hint: what happens if you have many values in `temp`?)*

```{r 3}
how_temp <- function(temp) {
  cut(
    temp,
    breaks = c(-Inf, 0, 10, 20, 30, Inf),
    labels = c("freezing", "cold", "cool", "warm", "hot"),
    right = TRUE
  )
}

how_temp(11)
how_temp(30)
how_temp(c(-2, -5, 4, 12, 45))
```
<p class="answer">
  <=가아니라 < 이면 cut 함수의 right 인수를 FALSE로 바꾸면 경계값의 오른쪽을 포함하지 않아서 같은 의미가 된다.<br>
  단순히 주어진 if_else chunk는 단일 벡터를 받아 단일 벡터를 출력하지만 cut으로 범주형 변수를 이용해 구현하면 벡터의 요소가 여러개여도 각각의 요소에 대해서 대응되는 문자열을 출력할 수 있다.
</p>


### Textbook(1st) 21.2.1 problem 2

Eliminate the for loop in each of the following examples by taking advantage of an existing function that works with vectors:

```{r pr, eval = FALSE}
out <- ""
for (x in letters) {
  out <- stringr::str_c(out, x)
}

x <- sample(100)
sd <- 0
for (i in seq_along(x)) {
  sd <- sd + (x[i] - mean(x)) ^ 2
}
sd <- sqrt(sd / (length(x) - 1))

x <- runif(100)
out <- vector("numeric", length(x))
out[1] <- x[1]
for (i in 2:length(x)) {
  out[i] <- out[i - 1] + x[i]
}
```

```{r 4}
#1
str_c(letters, collapse = "")
str_flatten(letters)

#2
x <- sample(100)
sd <- sqrt(sum((x - mean(x))^2) / (length(x) - 1))
sd

#3
x <- runif(100)
cumsum(x)
```

### Textbook(1st) 21.4.1 problem 1

Read the documentation for `apply()`. In the 2d case, what two for loops does it generalise?

<p class="answer">
  행별로 돌리는 for루프와 열별로 돌리는 for루프가 일반화 된다.
</p>

### Textbook(1st) 21.5.3 problem 2

How can you create a single vector that for each column in a data frame indicates whether or not it’s a factor?

```{r 5}
map_lgl(mtcars, is.factor)
map_lgl(diamonds, is.factor)
```

### Textbook(1st) 23.2.1 problem 2

One way to make linear models more robust is to use a different distance measure.<br>
For example, instead of root-mean-squared distance, you could use mean-absolute distance:

```{r given_function}
measure_distance <- function(mod, data) {
  diff <- data$y - (mod[1] + data$x * mod[2])
  mean(abs(diff))
}
```

Use optim() to fit this model to the simulated data above and compare it to the linear model.

```{r 6}
library(modelr)

measure_distance <- function(mod, data) {
  diff <- data$y - (mod[1] + data$x * mod[2])
  mean(abs(diff))
}

best <- optim(c(0, 0), measure_distance, data = sim1)
best$par


ggplot(sim1, aes(x, y)) +
  geom_point(size = 2, colour = "grey30") +
  geom_abline(intercept = best$par[1], slope = best$par[2])
```

<p class="answer">
  measure_distance를 root-mean-squared distance로 했을 때의 회귀계수 4.222248 2.051204와 다르게
  4.364852 2.048918이 나왔다.
  그래프를 보면 아웃라이어에 더 둔감하게 반응한다.
</p>

### Textbook(1st) 23.3.3 problem 1

Instead of using `lm()` to fit a straight line, you can use `loess()` to fit a smooth curve. Repeat the process of model fitting, grid generation, predictions, and visualisation on `sim1` using `loess()` instead of `lm()`. How does the result compare to `geom_smooth()`?

```{r 7}
grid <- sim1 %>%
  data_grid(x)

lmod <- loess(y ~ x, data = sim1)
pred <- grid %>%
  add_predictions(lmod)

ggplot(sim1, aes(x = x, y = y)) +
  geom_point(aes(y = y)) +
  geom_line(aes(y = pred), data = pred, colour = "red", size = 1) +
  geom_smooth(color = "blue", se = FALSE)
```

<p class="answer">
  빨간 선(직접 모델링)과 파란 선(geom_smooth)는 거의 겹치는 것을 알 수 있다.
</p>

### Textbook(1st) 23.4.5 problem 3
Using the basic principles, convert the formulas in the following two models into functions. *(Hint: start by converting the categorical variable into 0-1 variables.)*

```{r model, eval = FALSE}
mod1 <- lm(y ~ x1 + x2, data = sim3)
mod2 <- lm(y ~ x1 * x2, data = sim3)
```

```{r 8}
# mod1
x <- model.matrix(y ~ x1 + x2, sim3)
y = sim3$y
beta1 <- solve(t(x) %*% x) %*% t(x) %*% y
beta1

# mod2
x2 <- model.matrix(y ~ x1 * x2, sim3)
beta2 <- solve(t(x2) %*% x2) %*% t(x2) %*% y
beta2
```

### Textbook(2nd) 26.2.8 problem 1

Practice your `across()` skills by:

1. Computing the number of unique values in each column of `palmerpenguins::penguins`.

```{r 9.1}
palmerpenguins::penguins %>%
  summarise(across(everything(), ~length(unique(.))))
```

2. Computing the mean of every column in `mtcars`.

```{r 9.2}
mtcars %>%
  summarise(across(everything(), mean))
```

3. Grouping `diamonds` by `cut`, `clarity`, and `color` then counting the number of observations and computing the mean of each numeric column.

```{r 9.3}
diamonds %>%
  group_by(cut, clarity, color) %>%
  summarise(
    n = n(),
    across(where(is.numeric), mean)
  )
```
